---
title: "Stat 184 Final Project"
author: "Kelly Zhang, Tyler Wong, Yuheng Xie"
date: "`r paste('Last Updated:', Sys.Date())`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

## Introduction

**Are Used Car (Ford) Prices Different in Inner and Outer Cities?**

Since 2020, notably after the pandemic, car prices have soared to new heights. In response to economic inflation, specifically fuel prices, there’s no doubt that purchasing a new car requires considerable reflection and a strong financial position. One solution to consider is purchasing a used car. 

When it comes to purchasing a used car, many factors are taken into consideration. These include the price of the car, its mileage, the model, and the year. These elements have proven to be statistically significant, through numerous published experiments, in determining the price of a used car. However, have we considered the impact of the location where the car is purchased? 

For instance, the supply and demand in inner cities, such as New York or Philadelphia, may be higher than in rural or suburban areas like Harrisburg or Albany. Inner cities, with their larger populations, may have higher demand, leading to car price inflation. On the other hand, less populated areas may have less demand and could offer cheaper options. Therefore, a potential research question could be, “Are Used Car Prices Different in Inner and Outer Cities?”

To answer this question, our group used data from Autotrader. The Autotrader generator is a website that produces CSV datasets of used cars listed at autotrader.com, an online auto retailer, based on a maker, a model, and a zip code. In the dataset, each case represents an individual Ford car. We chose Ford as the focus model as it is the most popular car brand in the United States, allowing for a large amount of data and lower standard error. We randomly selected one major city and a suburban or rural area for comparison to allow for possible generalization. These locations are New York City and Albany, and Philadelphia and Harrisburg. The dataset includes the variables year, price, and mileage. To add location, we need to add a column titled “location.” This will be further explained in the data visualization preparation process.


## Data Visualization Preparation & Coding Style

**Coding Style**

The coding style used throughout this report is Book of Apps for Statistics Teaching (BOAST) Style.

**Data Visualization Preparation Process**

The research is divided into four parts, with the first part focusing on data wrangling and preparation. Because the data set is tidied, we can immediately begin the wranging process. Given our focus on price and location, we add a location column and merge all Ford models to create a data set for New York City. This process is replicated for Albany, Philadelphia, and Harrisburg, resulting in four datasets that represent location and price of used Ford cars. Then, the data frames are merged into completion, titled “All_Cars.” Furthermore, each dataset is filtered to eliminate N/A values and instances where the price is 0, as prices cannot be $0.  

**Load Packages**

```{r, include = F}
rm(list = ls()) #clear the workspace

library(knitr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggformula)
library(mosaic)
library(car)
library(janitor)
library(kableExtra)
library(psych)
library(tidyr)
library(tidyverse)

```

**Import Datasets**

```{r, include = F}
NYC_Taurus <- read.csv("Datasets/ALB_Edge.csv")
NYC_Mustang <- read.csv("Datasets/NY (Mustang).csv")
NYC_Fusion <- read.csv("Datasets/NY (Fusion).csv")
NYC_Focus <- read.csv("Datasets/NY (Focus).csv")
NYC_Fiesta <- read.csv("Datasets/NY (Fiesta).csv")
NYC_F350 <- read.csv("Datasets/NY (F350).csv")
NYC_F250 <- read.csv("Datasets/NY (F250).csv")
NYC_F150 <- read.csv("Datasets/NY (F150).csv")
NYC_Explorer <- read.csv("Datasets/NY (Explorer).csv")
NYC_Escape <- read.csv("Datasets/NY (Escape).csv")
NYC_Edge <- read.csv("Datasets/NY (Edge).csv")

ALB_Taurus <- read.csv("Datasets/ALB_Taurus.csv")
ALB_Mustang <- read.csv("Datasets/ALB_Mustang.csv")
ALB_Fusion <- read.csv("Datasets/ALB_Fusion.csv")
ALB_Focus <- read.csv("Datasets/ALB_Focus.csv")
ALB_Fiesta <- read.csv("Datasets/ALB_Fiesta.csv")
ALB_F350 <- read.csv("Datasets/ALB_F350.csv")
ALB_F250 <- read.csv("Datasets/ALB_F250.csv")
ALB_F150 <- read.csv("Datasets/ALB_F150.csv")
ALB_Explorer <- read.csv("Datasets/ALB_Explorer.csv")
ALB_Escape <- read.csv("Datasets/ALB_Escape.csv")
ALB_Edge <- read.csv("Datasets/ALB_Edge.csv")

PHIL_Taurus <- read.csv("Datasets/PHIL_Taurus.csv")
PHIL_Mustang <- read.csv("Datasets/PHIL_Mustang.csv")
PHIL_Fusion <- read.csv("Datasets/PHIL_Fusion.csv")
PHIL_Focus <- read.csv("Datasets/PHIL_Focus.csv")
PHIL_Fiesta <- read.csv("Datasets/PHIL_Fiesta.csv")
PHIL_F350 <- read.csv("Datasets/PHIL_F350.csv")
PHIL_F250 <- read.csv("Datasets/PHIL_F250.csv")
PHIL_F150 <- read.csv("Datasets/PHIL_F150.csv")
PHIL_Explorer <- read.csv("Datasets/PHIL_Explorer.csv")
PHIL_Escape <- read.csv("Datasets/PHIL_Escape.csv")
PHIL_Edge <- read.csv("Datasets/PHIL_Edge.csv")

HARS_Taurus <- read.csv("Datasets/HARS_Taurus.csv")
HARS_Mustang <- read.csv("Datasets/HARS_Mustang.csv")
HARS_Fusion <- read.csv("Datasets/HARS_Fusion.csv")
HARS_Focus <- read.csv("Datasets/HARS_Focus.csv")
HARS_Fiesta <- read.csv("Datasets/HARS_Fiesta.csv")
HARS_F350 <- read.csv("Datasets/HARS_F350.csv")
HARS_F250 <- read.csv("Datasets/HARS_F250.csv")
HARS_F150 <- read.csv("Datasets/HARS_F150.csv")
HARS_Explorer <- read.csv("Datasets/HARS_Explorer.csv")
HARS_Escape <- read.csv("Datasets/HARS_Escape.csv")
HARS_Edge <- read.csv("Datasets/HARS_Edge.csv")

```

**Add Location Column and Merge Datasets**

```{r}

NYC = rbind(NYC_Taurus, NYC_Mustang, NYC_Fusion, NYC_Focus, NYC_Fiesta, NYC_F350, NYC_F250, NYC_F150, NYC_Explorer, NYC_Escape, NYC_Edge) %>%
  select(-mileage) %>%
  filter(price != 0)

NYC$Location = rep("New York City", nrow(NYC))

ALB = rbind(ALB_Taurus, ALB_Mustang, ALB_Fusion, ALB_Focus, ALB_Fiesta, ALB_F350, ALB_F250, ALB_F150, ALB_Explorer, ALB_Escape, ALB_Edge) %>%
  select(-mileage) %>%
  filter(price != 0)

ALB$Location = rep("Albany", nrow(ALB))

PHIL = rbind(PHIL_Taurus, PHIL_Mustang, PHIL_Fusion, PHIL_Focus, PHIL_Fiesta, PHIL_F350, PHIL_F250, PHIL_F150, PHIL_Explorer, PHIL_Escape, PHIL_Edge) %>%
  select(-mileage) %>%
  filter(price != 0)

PHIL$Location = rep("Philadelphia", nrow(PHIL))

HARS = rbind(HARS_Taurus, HARS_Mustang, HARS_Fusion, HARS_Focus, HARS_Fiesta, HARS_F350, HARS_F250, HARS_F150, HARS_Explorer, HARS_Escape, HARS_Edge) %>%
  select(-mileage) %>%
  filter(price != 0)

HARS$Location = rep("Harrisburg", nrow(HARS))

NY <- 
  bind_rows(NYC, ALB) %>%
  na.omit()

PA <- 
  bind_rows(HARS, PHIL) %>%
  na.omit()

All_Cars <- 
  bind_rows(ALB, PHIL, NYC, HARS) %>%
  na.omit() 

write.csv(All_Cars, file = "All_Cars.csv") #add merged dataset to gitHUB

```

## Data Exploration and Visaulization

**Data Inspection**
```{r}
head(All_Cars)

glimpse(All_Cars)
```

Upon inspecting the data, all elements are correctly structured. The variables ‘year’ and ‘price’ are quantitatively measured as integers and floats, respectively, allowing for decimal values. As for ‘location’, it is a categorical variable represented by strings. Therefore, during Exploratory Data Analysis (EDA), we can anticipate the data frame to produce outputs for ‘price’ in decimal form and ‘location’ as strings. This is crucial as it enables us to generate certain graphs, such as boxplots, which account for one quantitative and one categorical variable.

**Side-by-Side Boxplot**

In the data exploration and visualization process, we create a boxplot using both the categorical and quantitative data. A box plot serves as an effective visualization tool to depict the relationship between these variables, as well as to display a five-number summary. 

***Five-number summary***: Descriptive statistics output the minimum value, first quartile (25th percentile), median (50th percentile), third quartile (75th percentile), and the maximum value.

##### Boxplot 1: New York

```{r}
ggplot(
data = NY,
mapping = aes(x = Location, y = price, fill = Location)) +
  geom_boxplot() +
  ggtitle("Boxplot of Used Ford Car Prices in New York") +
  scale_fill_manual(values = c("salmon", "salmon1")) +
  theme_bw() +
  xlab("Location") +
  ylab("Price (in USD)") +
  theme(
  legend.position = "none",
  text = element_text(size = 12))
```

##### Boxplot 2: Pennsylvania

```{r}
ggplot(
data = PA,
mapping = aes(x = Location, y = price, fill = Location)) +
  geom_boxplot() +
  ggtitle("Boxplot of Used Ford Car Prices in Pennsylvania") +
  theme_bw() +
  xlab("Location") +
  ylab("Price (in USD)") +
  scale_fill_manual(values = c("lightblue", "lightblue4")) +
  theme(
  legend.position = "none",
  text = element_text(size = 12))
```

**Boxplot Observations:**

In the initial comparative boxplot, which compares Albany and New York City, note the three features: the outliers, the interquartile range (IQR), and the median. During the data preparation phase, one of our objectives was to eliminate prices that were unrealistic in a real-world context, such as a price of $0 or NA. Outliers are challenging to eliminate, and given their abundance, it’s preferable to maintain these data points to preserve the majority of the dataset. Consequently, even though the maximum value is around $150,000, which may seem illogical compared to the median, we cannot determine whether or not it is a confirmed outlier. A significant advantage of a boxplot is its use of the median value instead of the mean. In most situations, the mean value would be preferred. However, due to the number of outliers, the median value provides a more accurate representation as it minimizes the impact of outliers, thereby preventing the influence of a potential skew.

The IQR (third quartile minus the first quartile) is similar to the median, suggesting that the distribution or spread of the data is close. This also implies that the prices of individual cars are likely to be similar.

The median is approximately $25,000 USD, and unfortunately, there appears to be insufficient evidence to suggest a correlation between the location of the car purchase and its price.

The second boxplot yields almost identical results to the first. The outliers appear to dominate the plot. Interestingly, the outliers and IQR closely match those of the New York plot, indicating that the prices of used Ford models are consistent across all locations, as are their medians. This observation will be further investigated in the forthcoming bar graph visualization.

### **Bar Graphs**

##### Bar Graph 1: New York

```{r}

NY_Median <- 
  NY %>%
  group_by(Location) %>%
  summarise(median_price = median(price), na.rm = TRUE)

ggplot(
data = NY_Median,
mapping = aes(x = Location, y = median_price, fill = Location)) +
  scale_fill_manual(values = c("salmon", "salmon1")) +
  geom_bar(stat = "identity", width = 0.3) +
  ggtitle("Bargraph of Used Ford Car Prices by Location") +
  theme_bw() +
  xlab("Location") +
  ylab("Median Price (USD)") +
  theme(
  legend.position = "none",
  text = element_text(size = 10))

```

##### Bar Graph 2: Pennsylvania

```{r}
PA_Median <- 
  PA %>%
  group_by(Location) %>%
  summarise(median_price = median(price), na.rm = TRUE)

ggplot(
data = PA_Median,
mapping = aes(x = Location, y = median_price, fill = Location)) +
  scale_fill_manual(values = c("lightblue", "lightblue4")) +
  geom_bar(stat = "identity", width = 0.3) +
  ggtitle("Bargraph of Used Ford Car Prices by Location") +
  theme_bw() +
  xlab("Location") +
  ylab("Median Price (USD)") +
  theme(
  legend.position = "none",
  text = element_text(size = 10))

```
**Bar Graph Observations:**

From the side-by-side boxplot, we concluded that there is a minor difference between the inner and outer city car prices. The bar graphs allow us to “zoom in” on this visualization by wrangling both data sets with the group_by() and summarise() functions. Using these functions, we can calculate the median for both data sets.

In the first bar graph, between Albany and New York City, the median is close to a margin of $2000 USD. From a numerical standpoint, a difference of $2000 is not significant when compared to the range of values, which is from around $1000 to $150,000 USD. However, in a real-life context, $2000 might be considered a deal breaker for most clients. Thus, it’s difficult to determine whether or not this difference is important based purely on EDA, without further hypothesis testing. But, with the result of the second graph and boxplot, we can create a strong inference.

When considering the median prices in Pennsylvania, Harrisburg and Philadelphia, the median is approximately $1000 USD. One similarity is that the average car prices in Harrisburg and Albany are higher than those in Philadelphia and New York City. On one hand, this is surprising as cities often struggle with economic inflation. On the other, outer cities may have a lower supply as there are fewer residents.

Based on the observations of both graphs, we remain confident that there is still no significant difference in price based on location.

### **Scatterplot Graphs**

##### Scatterplot Graph 1: New York

**Scatterplot Process & Purpose:**

The scatterplot represents a different approach from our general data visualization process. One of the main differences is that the scatterplot includes the variable 'year' on the x-axis, which means there are three variables of interest (price, year, location) instead of the original two (price and location). The scatterplot was introduced to display the overall trend of Ford car prices and to provide a visual aid for the two regression lines. Without each individual data point, it might be difficult to understand where the intersection between the lines occurs or why there is a positive slope.

```{r}

gf_point(price ~ year, color = ~ Location, data = NY) %>%
  gf_lm()

```

**Scatterplot 1 Observations**

In the first scatterplot, there are multiple observations to note. Firstly, it’s noticeable that there is an intersection point between the lines. This is interesting as it suggests that there might be a significant interaction term between location and price. Unfortunately, a hypothesis test would be required to make this inference.

Secondly, there is a noticeable increasing trend towards the year 2020. The values, almost in the shape of a helix, on the far right indicate an all-time-high price at around $150,000. Although this might be an outlier, it’s clear from the following data that the price increases as the year progresses. In the context of this data set, it provides evidence that the price of a car increases with its novelty.

Although this observation doesn’t necessarily support our research inquiry into the relationship between location and price, it underscores the importance of exploring other variables when the significance of location and price appears to be minimal. 


##### Scatterplot Graph 2: Pennsylvania

```{r}

gf_point(price ~ year, color = ~ Location, data = PA) %>%
  gf_lm()
```

**Scatterplot 2 Observations:**

The second scatterplot, which compares Harrisburg and Philadelphia, is notable because the regression lines are extremely similar. This contradicts the information from the first scatterplot, which showed small differences in slope and y-intercept, specifically between the years 1990 to 2010. This suggested that there might be a difference in car prices between locations. However, the scatterplot for Pennsylvania shows little evidence of a significant difference in car prices between Harrisburg and Philadelphia. 


### **Summary Table**

##### Summary Table 1: New York

```{r}
# Get descriptive statistics of price and location
groupStats_NY <- 
  psych::describeBy(
  x = NY$price,
  group = NY$Location,
  na.rm = TRUE,
  skew = TRUE,
  ranges = TRUE,
  quant = c(0.25, 0.75),
  IQR = TRUE,
  mat = TRUE,
  digits = 4)

# Set row names as location; select useful columns
groupStats_NY <- 
  groupStats_NY %>%
  tibble::remove_rownames() %>%
  tibble::column_to_rownames(
  var = "group1") %>%
  dplyr::select(
  n, min, Q0.25, median, Q0.75, max, mad, mean, sd, skew, kurtosis)

# Generate a professional looking table
groupStats_NY %>%
  knitr::kable(
    caption = "Summary Statistics for Used Car Prices (in Thousands) in New York",
    digits = 3,
    format.args = list(big.mark = ","),
    align = rep('c', 11),
    col.names = c("n", "Min", "Q1", "Median", "Q3", "Max", "MAD", "SAM", "SASD",
    "Sample Skew", "Sample Ex. Kurtosis"),
  booktabs = TRUE) %>%
  kableExtra::kable_styling(
  font_size = 12,
  latex_options = c("scale_down", "HOLD_position"))

```

**Summary Table 1 Observations:**


In the summary table, we are provided with multiple outputs, a five-number summary and informaton about the error. The **sample minimum** and **sample maximum** informs us about the lowest prices of each location. In the state of New York, the lowest prices for a used Ford is 1000 dollars USD in both the outer and inner city, and the highest price is approximately 130 thousand dollars in the outer city and 150 thousand dollars in the inner city. While this does not provide useful evidence towards our research question, it is informative in determining possible outliers within the minimum and maximum values.

The first quartile (Q1), or 25% of the data, lies within 20 thousand dollars, respectfully for both Albany and New York City, with a difference of 3000 dollars. The third quartile (Q3), or 75% of the data, are less than 38 thousand dollars and 37.45 thousand dollars. This is the higher end of the prices, but in comparison to the first quartile, the price differences are similar, with a simple 0.546 thousand distinction. In terms of MAD, SAM, SASD, and Sample Skew, these are useful in determining the overall distribution and error rate of the data. Notice, the standard deviation is quite high. This may be an issue, however, when accounting for the minimum and maximum values, ranging from 16 thousand dollars to 150 thousand dollars, a standard deviation of 15 to 16 thousand is understandable. 

The most significant aspect of the summary table is median. Because we are interested in determining an overall trend, the average (or median), can be useful in determining whether or not there is a significant difference in the car prices of all models. In New York, the median prices for inner and outer city are quite similar, with a 2.695 thousand dollar margin, while accounting for the range of values; however, to most, 2.695 thousand dollars would likely be a significant difference. 


##### Summary Table 2: Pennsylvania

```{r}
groupStats_PA <- 
  psych::describeBy(
  x = PA$price,
  group = PA$Location,
  na.rm = TRUE,
  skew = TRUE,
  ranges = TRUE,
  quant = c(0.25, 0.75),
  IQR = TRUE,
  mat = TRUE,
  digits = 4)

# Set row names as location; select useful columns
groupStats_PA <- 
  groupStats_PA %>%
  tibble::remove_rownames() %>%
  tibble::column_to_rownames(
  var = "group1") %>%
  dplyr::select(
  n, min, Q0.25, median, Q0.75, max, mad, mean, sd, skew, kurtosis)

# Generate a professional looking table
groupStats_PA %>%
  knitr::kable(
    caption = "Summary Statistics for Used Car Prices (in Thousands) in Pennsylvania",
    digits = 3,
    format.args = list(big.mark = ","),
    align = rep('c', 11),
    col.names = c("n", "Min", "Q1", "Median", "Q3", "Max", "MAD", "SAM", "SASD",
    "Sample Skew", "Sample Ex. Kurtosis"),
  booktabs = TRUE) %>%
  kableExtra::kable_styling(
  font_size = 12,
  latex_options = c("scale_down", "HOLD_position"))

```

**Summary Table 2 Observations**

In table 2, the minimum values are 2 thousand dollars and 1.795 thousand dollars. Interestingly, this is higher than the comparison in New York. While this does not satisfy our research question, it could provide insight towards prices differing based on state. Interestingly, the maximum value are the same in both areas, 148 thousand dollars. Because the average is around 25 thousand dollars, it is likely that, similar to the New York maximum, this data point could be a potential outlier. Thus, the maximum and minimum values are not entirely reliable for our research question.

Regarding the first quartile, Q1, 25% of the recorded prices are below 17 thousand dollars, respectively. There is a slim difference of 0.703 thousand dollars. The third quartile prices, Q3, are less than 37.9 thousand dollars and 36.988 thousand dollars. Again, the difference is not significant enough to state that there is significance in price based on location. From the summary table of Pennsylvania, the prices are higher as whole, but when compared to the summary of New York, there is less of a location difference. This is mildly contradicting to our research question; however, it is insight to consider other confounding variables that may arise from the data set, such as year of the model.

The median of the locations, 25.769 and 24.995 thousand dollars further highlight the average price of a Used Ford car to be around mid 20 thousand dollars, whether or not it is in the inner or outer city, when using evidence from both Pennsylvania and New York. There does not seem to be a significant difference in the median values, and based on the full analysis of the table, including the five-number summary and maximum and minimum, Pennsylvania does not have any important distinction in where consumers should purchase used Fords. 

### **Conclusion**

**Research Question: Are Used (Ford) Car Prices Different in Inner and Outer Cities?**

In conclusion, based on the EDA and various visualizations presented, we did not find evidence that there is a statistically significant difference between used Ford car prices in inner and outer cities. Although there was no hypothesis test, it can be predicted that the p-value would be larger than a 5% threshold. To summarize, the boxplot provided us with the distribution, spread, and general possibility of outliers in the data set. The IQR for Pennsylvania and New York did not differ much between inner and outer cities. To further emphasize this fact, we used a bargraph with the median as the response variable, to "zoom in" on the average differences. We found that there are minor distinctions between the median values. Next, to explore the variable year, and provide a general visualization of each individual case, along with a regression line to portray the upward trend, we used a scatterplot. While there was nothing new about the median, we discovered that there was an interaction between the two regression lines, and with further anaylsis, such as a hypothesis test between interactive terms, there may be a significant p-value. Finally, with the summary tables, we were able to discuss the numerical differences and found that 2 thousand dollars was the largest/most significant found. Overall, it can be assumed that location is not a significant variable when considering used car prices based on the Ford model.


